% Abstract on the report fot the NXP S32K3X8EVB board on QEMU
\chapter{Abstract}
\label{Abstract}

This report presents the development and implementation of a simulation environment for the NXP S32K3X8EVB board using QEMU. The primary objective of this project is to facilitate the testing and development of embedded software applications without the need for physical hardware, thereby reducing costs and increasing accessibility for developers.
The report begins with an overview of the NXP S32K3X8EVB board, highlighting its key features and specifications. It then delves into the architecture and capabilities of QEMU, emphasizing its suitability for simulating embedded systems. The core of the report focuses on the step-by-step process of setting up the QEMU environment to emulate the S32K3X8EVB board, including the configuration of necessary peripherals and interfaces (UART and SPI).
Furthermore, the report discusses the challenges encountered during the implementation phase and the solutions devised to overcome them.

Finally, the report concludes with an evaluation of the simulation environment's performance and its potential applications in embedded systems development. The successful implementation of this project demonstrates the feasibility of using QEMU for simulating complex embedded systems, paving the way for future advancements in this field.
The report is structured to provide a comprehensive understanding of the project, making it a valuable resource for developers and researchers interested in embedded systems simulation. The structure is as follows:
\begin{itemize}
    \item Chapter 1: Introduction - Provides background information and outlines the objectives of the project.
    \item Chapter 2: NXP S32K3X8EVB Board Overview - Details the specifications and features of the board.
    \item Chapter 3: QEMU Overview - Discusses the architecture and capabilities of QEMU.
    \item Chapter 4: Implementation - Describes the step-by-step process of setting up the simulation environment.
    \item Chapter 5: FreeRTOS Porting - Explains the process of porting FreeRTOS to the simulated environment.
    \item Chapter 6: Testing and Conclusion - Evaluates the performance of the simulation environment and summarizes key findings.
    \item References - Lists the sources and references used throughout the report.
    \item Appendices - Includes supplementary material such as code snippets and configuration files.
\end{itemize} 

A prior knowledge of embedded systems, real-time operating systems, transmission protocols and virtualization technologies is beneficial for understanding the content of this report.